The Talking Heads Experiment

### 2015-04-17 chap10/chap10.tex
orphans, widows, line breaks

From 1996, even before the first Talking Heads experiment was started, experiments were already conducted in our 
laboratory, primarily by Paul Vogt, exploring language games on the cybernetic mobile robots that we were able to build ourselves
at the time. These robots were constructed from Lego bricks and 
used a basic processing board for linking directly simple sensors (touch sensors, infrared sensors) and actuators (left and right motors)
using an adaptive dynamical system (see \figref{f:plate3}). They were however too unreliable for long-term repeatable 
experimentation and the sensori-motor experiences were too restricted to hope for the development of interesting languages. The use 
of pan-tilt cameras, as in the Talking Heads experiment, was an attempt to 
have an experimental set-up at relatively low cost that was reliable and
used vision as the source of information about the world. Of course this came
at the price of less mobility and no true physical interaction with the real world. Nevertheless, many further fruitful experiments were 
done using the same set-up, particularly to explore in much greater depth the domain of color lexicons
\footnote{Colour became a focal point of research because there is an extensive literature in cognitive science that has 
been gathering empirical data about color and its evolution and because it is relatively straightforward to do 
colour naming games. One of the main papers on our colour naming research is 
\cite{Steels:2005}.} and multi-word naming games\footnote{Multi-word games were the focus of the Ph.D research by 
Joris Van Looveren, with a thesis defended in 2004.}
A Talking Heads simulator was also built by Paul Vogt in 
order to speed up such experiments and prepare the way for more advances with real robots \citep{Vogt:2003}.

>>>>

From 1996, even before the first Talking Heads experiment was started, experiments were already conducted in our 
laboratory, primarily by Paul Vogt, exploring language games on the cybernetic mobile robots that we were able to build ourselves
at the time. These robots were constructed from Lego bricks and 
used a basic processing board for linking directly simple sensors (touch sensors, infrared sensors) and actuators (left and right motors)
using an adaptive dynamical system (see \figref{f:plate3}). They were however too unreliable for long-term repeatable 
experimentation and the sensori-motor experiences were too restricted to hope for the development of interesting languages. The use 
of pan-tilt cameras, as in the Talking Heads experiment, was an attempt to 
have an experimental set-up at relatively low cost that was reliable and
used vision as the source of information about the world. Of course this came
at the price of less mobility and no true physical interaction with the real world. Nevertheless, many further fruitful experiments were 
done using the same set-up, particularly to explore in much greater depth the domain of color lexicons\footnote{Colour became a focal point of research because there is an extensive literature in cognitive science that has 
been gathering empirical data about color and its evolution and because it is relatively straightforward to do 
colour naming games. One of the main papers on our colour naming research is 
\cite{Steels:2005}.} and multi-word naming games.\footnote{Multi-word games were the focus of the Ph.D research by 
Joris Van Looveren, with a thesis defended in 2004.}
A Talking Heads simulator was also built by Paul Vogt in 
order to speed up such experiments and prepare the way for more advances with real robots \citep{Vogt:2003}.

### 2015-04-17 chap11/chap11.tex
orphans, widows, line breaks

The general set-up for the Proper Naming Game experiment
is shown in \figref{fig:grounded}. We see two \textsc{qrio} robots and an environment consisting of objects of various 
sizes and shapes. At any time a new object can be added or the position of an object can be changed. So the world is 
entirely open and robots do not know how many individuals there are. 
Importantly, the robots had to recognise an object as an individual, independently of its position. If 
a specific object disappeared from the world and was put back later, the same name would still have to be used. \\

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{chap10/figs/prototypes-identifiers-words.pdf}}
\caption{\label{fig:prototypes-identifiers-words}In the Proper Naming Game experiment the memory of the agents is conceived in terms of a semiotic network that 
spans the full semiotic cycle from sensory experience and conceptualisation to naming.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{chap11/figs/png-results.pdf}}
\caption{\label{fig:png-results} 
{\itshape Left:} Results for Proper Naming games in a population of 10 agents that have to name 15 individual objects. 
We see that a vocabulary establishes
itself after agents have played a few hundred games. There is again the typical peak until alignment kicks in. 
The number of individuals is larger than 15, implying that agents name prototypical views rather than 
individuals. {\itshape Right:} Results after the addition  
of a heuristic based on tracking an object over time. We see that there are fewer individuals 
than stored prototypical views, which means that agents have improved their ability to recognise object 
identity. }
\end{figure}
With these mechanisms, the population reaches a high level of communicative success (above 90 \%). 
However the average number of individuals in the agents' semiotic networks
is much larger than the number of distinctive objects
introduced in the experiment (see \figref{fig:png-results}, left). Apparently, agents are naming prototypical views of
individual objects instead of the individuals themselves. So we did not adequately address 
how a robot learns to interrelate different views. To do this, various heuristics need to be added. 
\clearpage
We have operationalised just one example of such a heuristic. When an object is moving or being moved,
its appearance may change but the observer tracking the object 
still knows that he is dealing with the same object and so he can
exploit this information to fine-tune his semiotic networks. Because the vision system we used on the \textsc{qrio} robots, developed 
by Michael Spranger,\footnote{
This system is described in-depth in: \cite{Spranger:2012vision}.} is able to track an object, it can align object
regions across different images, enriched with top-down predictions of
how an object region will change over time based on Recursive Bayesian
estimation using Kalman filters. The object being moved by 
the experimenter in \figref{fig:png-perception} 
yields two quite different sensory experiences when standing up or lying down which match with 
two different prototypes, but thanks to the tracking heuristic, the semiotic network can 
be rearranged to reflect that they are two views of the same object (see \figref{fig:png-optimisation}).

@@

A language strategy has two components: (i) a particular way of conceptualising reality (and learning the categories involved
in this conceptualisation) and (ii) a particular grammatical pattern that suggests to the listener what strategy is intended 
and provides the contents to execute the strategy. For example, in the phrase `very green', the combination of an adverb 
`very' and a basic colour adjective `green' suggests a ``graded membership strategy'' where not only a basic colour prototype
(`green') is introduced but also how close the sample is to this prototype (`very'). I will now introduce first how 
the conceptual aspect of a language strategy is defined using a procedural semantics system called IRL and then how the 
linguistic aspect is defined, using Fluid Construction Grammar. \\

@@

Constraint networks are an obvious choice for grounded (natural) language processing. \is{constraint network}
For example, to interpret the sentence `the ball behind
the block', it is possible that there are many blocks, many balls, but only one ball that is behind a block. So the 
unique referent of `the ball', `the block' and of the total noun phrase only become clear when 
all constraints have been applied. Natural languages clearly do not supply control information but assume that 
the interpreter is smart enough to interpret constraints based on data flow. The other properties of constraint 
programming are also important. The constraint network planner can be used for conceptualising what to say and 
the constraint network matcher is highly useful in parsing an incomplete language fragment or in learning the 
meaning of a new grammatical construction. \\

@@

The constraint network planner works on an evolutionary basis, similar to genetic algorithms. Starting from the goal 
(for example identifying a particular object in the environment), it will generate possible networks by recruiting 
existing primitive operations, linking them in, and checking whether the computation makes progress towards reaching the goal. 
The slots of a constraint are typed to limit what semantic operations will be tried next. Networks that perform useful 
functions are ``chunked'' into higher order units that then can be used as units in their own right, thus scaling up 
complexity. These chunked networks become associated with stereotyped grammatical expression, so that there is a very 
fast retrieval and semantic execution. \\

\subsection{Strategies for colour}
% {\bfshape Strategies for Colour} \\

Let us now look at concrete examples from the colour domain as worked out by Joris Bleys, studying the following 
strategies: \is{colour strategies}

1. The first strategy is the {\bfshape Basic Colour Strategy} which implements a prototype-based approach to 
the basic colour categories. This approach is similar to the one used in the Proper Naming Game, except that 
only the three colour-dimensions are profiled (the opponent channels red-green, yellow-blue and the brightness channel). 
The strategy is similar to the one shown earlier. 
\begin{verbatim}
(equal-to-context ?s1) 
(profile-colour-dimensions ?s2 ?s1)
(filter-by-colour-lenient ?s3 ?s2 ?cc)
(select-most-activated ?t ?s3)
\end{verbatim}
When the speaker says `green', this network is used together with a specific binding of ?cc to the prototype for 
green (denoted as [green]). 
\begin{verbatim}
(bind colour-category ?cc [green])
\end{verbatim}
A graphical representation of this network is shown in \figref{fig:basic-strat}.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/graded.pdf}}
\caption{\label{fig:graded}Constraint network implementing the Graded Membership Strategy. It builds further on the network for Basic Colour
Categorisation but adds a constraint to handle graded membership. }
\end{figure}

The network is entirely general for all expressions that use graded membership and gets instantiated for a particular 
case by specifying the semantic entities involved using the bind operation. For example, Tarahumara has the word 
`sita' for a particular colour prototype and `kame' for one type of graded membership: 
\begin{verbatim}
(bind colour-category ?cc [sita])
(bind membership-category ?mc [kame]) 
\end{verbatim}

@@

Many other strategies for colour can be imagined and some of them are found in human languages (although not necessarily in English). 
These strategies always involve operations over the colour space (such as rescaling, dimensionality reduction, etc.) alongside 
additional operations (such as getting the context, picking out a referent, computing the intersection of two sets, etc.). \\

@@

Functional constructions create a unit on top of the word. For example, in the Basic Colour Strategy, 
the colour-category is used in the filter-by-colour-lenient operation (see \figref{fig:basic-strat}). 
The new unit is called colour-unit in \figref{fig:syntactic-graded}, left. 
The meaning specifies the operations to be added to the network, which is in this case: 
\begin{verbatim}
((filter-by-colour-lenient ?s3 ?s2 ?c))
\end{verbatim}
\clearpage
There are two variables that can be linked to other operations with this meaning: ?s3 and ?s2 both of type object-set. 
On the syntactic side, the unit is categorised as a constituent named colour-category.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap11/figs/grounding-comparison.pdf}}
\caption{Comparison of three different environmental set-ups (left, right, and middle columns)
and four different learning conditions. Communicative success is shown on the y-axis.}
\label{fig:grounding-comparison} 
\end{figure}

>>>>

The general set-up for the Proper Naming Game experiment
is shown in \figref{fig:grounded}. We see two \textsc{qrio} robots and an environment consisting of objects of various 
sizes and shapes. At any time a new object can be added or the position of an object can be changed. So the world is 
entirely open and robots do not know how many individuals there are. 
Importantly, the robots had to recognise an object as an individual, independently of its position. If 
a specific object disappeared from the world and was put back later, the same name would still have to be used.

@@

\begin{figure}[b]
  \centerline{\includegraphics[width=.95\textwidth]{chap10/figs/prototypes-identifiers-words.pdf}}
\caption{\label{fig:prototypes-identifiers-words}In the Proper Naming Game experiment the memory of the agents is conceived in terms of a semiotic network that 
spans the full semiotic cycle from sensory experience and conceptualisation to naming.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap11/figs/png-results.pdf}}
\caption{\label{fig:png-results} 
{\itshape Left:} Results for Proper Naming games in a population of 10 agents that have to name 15 individual objects. 
We see that a vocabulary establishes
itself after agents have played a few hundred games. There is again the typical peak until alignment kicks in. 
The number of individuals is larger than 15, implying that agents name prototypical views rather than 
individuals. {\itshape Right:} Results after the addition  
of a heuristic based on tracking an object over time. We see that there are fewer individuals 
than stored prototypical views, which means that agents have improved their ability to recognise object 
identity. }
\end{figure}
With these mechanisms, the population reaches a high level of communicative success (above 90 \%). 
However the average number of individuals in the agents' semiotic networks
is much larger than the number of distinctive objects
introduced in the experiment (see \figref{fig:png-results}, left). Apparently, agents are naming prototypical views of
individual objects instead of the individuals themselves. So we did not adequately address 
how a robot learns to interrelate different views. To do this, various heuristics need to be added. 
\clearpage
We have operationalised just one example of such a heuristic. When an object is moving or being moved,
its appearance may change but the observer tracking the object 
still knows that he is dealing with the same object and so he can
exploit this information to fine-tune his semiotic networks. Because the vision system we used on the \textsc{qrio} robots, developed 
by Michael Spranger,\footnote{
This system is described in-depth in: \cite{Spranger:2012vision}.} is able to track an object, it can align object
regions across different images, enriched with top-down predictions of
how an object region will change over time based on Recursive Bayesian
estimation using Kalman filters. The object being moved by 
the experimenter in \figref{fig:png-perception} 
yields two quite different sensory experiences when standing up or lying down which match with 
two different prototypes, but thanks to the tracking heuristic, the semiotic network can 
be rearranged to reflect that they are two views of the same object (see \figref{fig:png-optimisation}).

@@

A language strategy has two components: (i) a particular way of conceptualising reality (and learning the categories involved
in this conceptualisation) and (ii) a particular grammatical pattern that suggests to the listener what strategy is intended 
and provides the contents to execute the strategy. For example, in the phrase `very green', the combination of an adverb 
`very' and a basic colour adjective `green' suggests a ``graded membership strategy'' where not only a basic colour prototype
(`green') is introduced but also how close the sample is to this prototype (`very'). I will now introduce first how 
the conceptual aspect of a language strategy is defined using a procedural semantics system called IRL and then how the 
linguistic aspect is defined, using Fluid Construction Grammar.

@@

Constraint networks are an obvious choice for grounded (natural) language processing. \is{constraint network}
For example, to interpret the sentence `the ball behind
the block', it is possible that there are many blocks, many balls, but only one ball that is behind a block. So the 
unique referent of `the ball', `the block' and of the total noun phrase only become clear when 
all constraints have been applied. Natural languages clearly do not supply control information but assume that 
the interpreter is smart enough to interpret constraints based on data flow. The other properties of constraint 
programming are also important. The constraint network planner can be used for conceptualising what to say and 
the constraint network matcher is highly useful in parsing an incomplete language fragment or in learning the 
meaning of a new grammatical construction.

@@

The constraint network planner works on an evolutionary basis, similar to genetic algorithms. Starting from the goal 
(for example identifying a particular object in the environment), it will generate possible networks by recruiting 
existing primitive operations, linking them in, and checking whether the computation makes progress towards reaching the goal. 
The slots of a constraint are typed to limit what semantic operations will be tried next. Networks that perform useful 
functions are ``chunked'' into higher order units that then can be used as units in their own right, thus scaling up 
complexity. These chunked networks become associated with stereotyped grammatical expression, so that there is a very 
fast retrieval and semantic execution.
\subsection{Strategies for colour}
% {\bfshape Strategies for Colour} \\

Let us now look at concrete examples from the colour domain as worked out by Joris Bleys, studying the following 
strategies: \is{colour strategies}
\enlargethispage{1\baselineskip}
1. The first strategy is the {\bfshape Basic Colour Strategy} which implements a prototype-based approach to 
the basic colour categories. This approach is similar to the one used in the Proper Naming Game, except that 
only the three colour-dimensions are profiled (the opponent channels red-green, yellow-blue and the brightness channel). 
The strategy is similar to the one shown earlier. 
\begin{verbatim}
(equal-to-context ?s1) 
(profile-colour-dimensions ?s2 ?s1)
(filter-by-colour-lenient ?s3 ?s2 ?cc)
(select-most-activated ?t ?s3)
\end{verbatim}
When the speaker says `green', this network is used together with a specific binding of ?cc to the prototype for 
green (denoted as [green]). 
\begin{verbatim}
(bind colour-category ?cc [green])
\end{verbatim}
A graphical representation of this network is shown in \figref{fig:basic-strat}.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{chap11/figs/graded.pdf}}
\caption{\label{fig:graded}Constraint network implementing the Graded Membership Strategy. It builds further on the network for Basic Colour
Categorisation but adds a constraint to handle graded membership. }
\end{figure}
\enlargethispage{1\baselineskip}
The network is entirely general for all expressions that use graded membership and gets instantiated for a particular 
case by specifying the semantic entities involved using the bind operation. For example, Tarahumara has the word 
`sita' for a particular colour prototype and `kame' for one type of graded membership: 
\begin{verbatim}
(bind colour-category ?cc [sita])
(bind membership-category ?mc [kame]) 
\end{verbatim}

@@

Many other strategies for colour can be imagined and some of them are found in human languages (although not necessarily in English). 
These strategies always involve operations over the colour space (such as rescaling, dimensionality reduction, etc.) alongside 
additional operations (such as getting the context, picking out a referent, computing the intersection of two sets, etc.).

@@

Functional constructions create a unit on top of the word. For example, in the Basic Colour Strategy, 
the colour-category is used in the filter-by-colour-lenient operation (see \figref{fig:basic-strat}). 
The new unit is called colour-unit in \figref{fig:syntactic-graded}, left. 
The meaning specifies the operations to be added to the network, which is in this case: 
\begin{verbatim}
((filter-by-colour-lenient ?s3 ?s2 ?c))
\end{verbatim}
There are two variables that can be linked to other operations with this meaning: ?s3 and ?s2 both of type object-set. 
On the syntactic side, the unit is categorised as a constituent named colour-category.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.9\textwidth]{chap11/figs/grounding-comparison.pdf}}
\caption{Comparison of three different environmental set-ups (left, right, and middle columns)
and four different learning conditions. Communicative success is shown on the y-axis.}
\label{fig:grounding-comparison} 
\end{figure}
